### 2020-05-07 README.Rmd
updated analysis structure and README

---
title: 'Terrebonne Project Summary'
author: 
date: 
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*insert terrebonne image here*

This document summarizes data within Terrebonne Bay, Louisiana (*lat. long.*) collected by Louisiana Universities Marine Consortium (LUMCON).

# Metadata summary

*data structure*

## Sampling locations map
*insert sampling locations, i.e. trawls, here*

# Environmental Data

A suite of environmental data have been collected through <a href="https://lumcon.edu/environmental-monitoring/" target="_blank">LUMCON's environmental monitoring program</a> over the course of the study period. These variables include:

- Water Temperature
- Salinity
- Specific Conductivity