### 2020-05-07 text/trawl_summary.Rmd

---
title: 'Trawl Summary'
author: "Jim Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
library(betapart)

TB_trawl_data <- readRDS("../data/TB_trawl_data.rds")
```

#TERREBONNE BAY TRAWL ANALYSIS SUMMARY {.unlisted .unnumbered}

*insert terrebonne image here*

This document summarizes trawl catch data within Terrebonne Bay, Louisiana (*lat. long.*) conducted by the Education Department of the Louisiana Universities Marine Consortium (LUMCON) from `r paste0(unique(min(TB_trawl_data$year))," - ",unique(max(TB_trawl_data$year)))`.

# Metadata summary

## Sampling locations map
*insert sampling locations, i.e. trawls, here*

# Environmental Data

A suite of environmental data have been collected through <a href="https://lumcon.edu/environmental-monitoring/" target="_blank">LUMCON's environmental monitoring program</a> over the course of the study period. These variables include:

- Water Temperature