### 2020-05-11 text/trawl_summary.Rmd
updated trawl_summary.RMD, added in docs/ folder for github.pages site.

---
title: 'Trawl Summary'
author: "Jim Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
library(betapart)
library(lubridate)

TB_trawl_data <- readRDS("../data/TB_trawl_data.rds")

```

#TERREBONNE BAY TRAWL ANALYSIS SUMMARY {.unlisted .unnumbered}

*insert terrebonne image here*

This document summarizes trawl catch data within Terrebonne Bay, Louisiana (*lat. long.*) conducted by the Education Department of the Louisiana Universities Marine Consortium (LUMCON) from `r paste0(unique(min(TB_trawl_data$year))," - ",unique(max(TB_trawl_data$year)))`.

# Metadata summary

## Sampling locations map
*insert sampling locations, i.e. trawls start and stop, here*

# Environmental Data

A suite of environmental data have been collected through <a href="https://lumcon.edu/environmental-monitoring/" target="_blank">LUMCON's environmental monitoring program</a> over the course of the study period. These variables include:

- Water Temperature
- Salinity
- Specific conductivity
- Dissolved Oxygen

```{r import environmental data, include = FALSE}

  temp_paths <- list.files(path = "../data/", "*temp_*", full.names = TRUE)
  temp_files <- lapply(temp_paths, read_csv, skip = 1,
                       col_types = cols(TS = col_datetime(format = "%m/%d/%Y %H:%M")))

  roundUP <- function(x,to=10){
    to*(x%/%to + as.logical(x%%to))
  }

  tempF_to_C <- function(F){(5/9)*(F-32)}

  temp_df <- bind_rows(temp_files) %>%
    rename(datetime = 'TS', temp_F = 'F') %>%
    mutate(yday = yday(datetime))

  do_paths <- list.files(path = "../data/", "*DO_*", full.names = TRUE)
  do_files <- lapply(do_paths, read_csv, skip = 1,
                     col_types = cols(TS = col_datetime(format = "%m/%d/%Y %H:%M")))

  do_df <- bind_rows(do_files) %>%
    rename(datetime = 'TS', do_mg_L = 'mg L') %>%
    mutate(yday = yday(datetime))

  hydro_paths <- list.files(path = "../data/", "hydro{1}.*.csv", full.names= TRUE)
  hydro_files <- lapply(hydro_paths, read_csv, col_type = cols(ArrayID = col_skip(), Year = col_skip(), JDay= col_skip(),
                                                               X4 = col_skip(), X5 = col_skip(), X6 = col_skip(), Time = col_skip(), X8 = col_skip(), `Date/Time` = col_character(),
                                                               DO = col_skip(), Value4 = col_double(), Depth = col_skip(), `Depth feet` = col_skip()))
  hydro_df <- bind_rows(hydro_files) %>%
    rename(datetime = 'Date/Time', temp_C = 'Temp', temp_F = 'Water Temp F', spCon_uS_cm = 'Value4', sal_psu = 'Sal') %>%
    mutate(datetime = parse_date_time(datetime, orders = c("%m/%d/%Y %H:%M", "%m/%d/%y %H:%M")),
           yday = yday(datetime))
env_long <- hydro_df %>% select(-yday, -temp_F) %>% bind_rows(temp_df %>% mutate(temp_C = tempF_to_C(temp_F)) %>% select(-yday, -temp_F)) %>%
  full_join(do_df %>% select(-yday)) %>%
  # full_join(temp_df %>% mutate(temp_C = tempF_to_C(temp_F))%>% select(-yday, - temp_F), all = TRUE) %>%
  pivot_longer(cols = temp_C:do_mg_L, names_to = "env_var", values_to = "value")

```

These data sets range in coverage, and are being added to continuously:

`r knitr::include_graphics("../figures/env_facet_plot.png")`

More data will be necessary to fully complete analyses, but this shows the emerging environmental picture.

# Biological Data

The education department has been conducting trawl surveys pretty consistently since 2007. These sampling dates are pretty evenly distributed throughout the year

## Sampling Events

```{r sampling events, include = TRUE, echo = FALSE}

TB_trawl_samplings <- TB_trawl_data %>% select(Date) %>%
  group_by(Date) %>% distinct %>% ungroup %>%
  mutate(Date = as.Date(Date), 
         Date_num = as.numeric(Date))

date_breaks = as.numeric(as.Date(paste0(seq(2007,2020, by = 1),"-01-01")))
date_labels = year(as.Date(date_breaks, origin = "1970-01-01"))

ggplot(TB_trawl_samplings, aes(x = Date, y = 0))+ geom_vline(aes(xintercept = as.numeric(ymd(Date)))) +
  scale_x_continuous(breaks = date_breaks, labels = date_labels) + theme_minimal() +
  theme(axis.title = element_blank())

```