### 2021-07-28 analyses/trawl/trawl_README.Rmd
large restructure of repository. Manuscript started.

---
title: "Trawl-README"
author: "JRJ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      out_dir <- "docs";
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(here::here(), out_dir, 'trawl_README.html')) })
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
here::i_am("analyses/trawl/trawl_README.Rmd")

```

This document outlines the workflow for the initial trawl analysis. The first manuscript's goals is to characterize and quantify the change in trawl communities in Terrebonne Bay from 2007 onward. Analysis sequence is:

1. "packages.R"
2. "datascript.R" (sourcing this document will call the 'packages.R' script)
3. "trawl_rarefaction.R"
4. "trawl_specaccum.R"