### 2022-09-28 sub-projects/trawl/docs/ms/ms.RMD
updated ms

---
title: "A decade of biodiversity change in an estuarine assemblage: fits and starts and natural events"
author:
   - name: James R Junker
     email: james.junker1@gmail.com
     institute: [glrc, lumcon]
     correspondence: true
   - name: Craig R McClain
     institute: lumcon
institute:
   - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
   - lumcon: Louisiana Universities Marine Consortium, Chauvin, LA USA
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

```{r Rmarkdown setup, include = FALSE, echo=F, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
here::i_am("sub-projects/trawl/docs/ms/ms.RMD")

source(here::here("datascript.R"))
source(here::here("sub-projects/trawl/R/Lorenz.R"))
source(here::here("sub-projects/trawl/R/Evenness.R"))

library(mgcv)
library(brms)
library(bayesplot)

rerunEven  = FALSE
rerunD = FALSE
# which years should we remove
year_removals = c("2007","2020","2021","2022")

# create the base data set
TB_trawl_data = TB_trawl_data %>%
   dplyr::filter(year %ni% year_removals)

TB_trawl_taxasite <- TB_trawl_data %>%
  dplyr::mutate(Date = as.Date(paste0(as.character(year),"-",as.character(month),"-01"))) %>%
  dplyr::select(Date, date_id, Common_name, Abundance) %>%
  group_by(Date, date_id, Common_name) %>%
  dplyr::summarise(Abundance = sum(Abundance)) %>%
  na.omit %>% group_by(Date, date_id, Common_name) %>%
  # # remove trawls with less than 10 individuals 
  # dplyr::filter(sum(Abundance) >=10) %>%
  # group_by(Date, date_id, Common_name) %>%
  pivot_wider(names_from = Common_name, values_from = Abundance, values_fill = list(Abundance = 0)) %>%
  ungroup() %>%
  dplyr::select(Date, date_id, everything())

TB_trawl_taxasite_long = TB_trawl_taxasite %>%
  pivot_longer(-Date:-date_id, names_to = "Common_name", values_to = "Abundance")

TB_trawl_taxasite_long %>%
  dplyr::filter(!duplicated(Common_name )) %>%
  dplyr::mutate(Date = as.Date(Date)) %>%
  arrange(Date) %>%
  dplyr::mutate(richness = 1:n())-> TB_time_accum

TB_trawl_effort_mth = TB_trawl_data %>%
  dplyr::filter(year(Date) %ni% year_removals) %>%
  dplyr::select(Date, date_id) %>%
  group_by(date_id) %>%
  dplyr::summarise(trawl_n = n()) %>%
  dplyr::mutate(Date = as.Date(date_id, origin = "2007-01-01")) %>%
  dplyr::select(Date, trawl_n)

# set specific variables used in methods
## how many bootstrap permutations to run with replacement? Keep low until final draft
nperm = 99
## how many individuals for count for Sn, Sn = 18 * n_year
n_mth = 2
#
iter = 7000
warmup = 5000
thin = 10
adapt_delta = 0.99

```
<!--- this is a note in Rmarkdown --->
\newpage

# Abstract

Anthropogenic activities continue to exert long-term, varied, and increasing pressures on marine ecosystems causing alterations to biological diversity on broad and local scales. Shifts in biodiversity are variable, nuanced and difficult to predict, yet, understanding how communities change through time is crucial for predicting current and future changes to biological systems and the services they provide. We document a decade of biodiversity change in the nekton community of a shallow-estuary...

\newpage

# Introduction

<!---  --->
Anthropogenic activities continue to exert long-term, varied, and increasing pressures on ecosystems [@steffen2015] causing alterations to biological diversity on local and global scales. Shifts in biodiversity at local scales are variable and potentially more nuanced than at global scales, but generally local changes manifest primarily as changes to community composition and structure rather than declines in richness [@hillebrand2018]. These shifts in communities through time are difficult to predict, in part, due to the multidimensional nature of biodiversity [@chase2018]. Yet, understanding how communities change through time is crucial for predicting current and future changes to biological systems and ecosystem services.

<!--- --->
Explaining the extent and pace of biodiversity change will require identifying the conditions that lead to divergent biological responses to environmental change, both natural and anthropogenic. In particular, it will be important to disentangle the roles of intrinsic (e.g., ecological) and extrinsic (e.g., environmental) drivers of biodiversity change. Intrinsic and extrinsic processes can interact, leading to temporal patterns of biodiversity change that vary in response to environmental change. For example, gradual environmental change over time can lead to gradual community shifts or these shifts may be undetectable or abrupt [@williams2011; @scheffer2003].

<!--- --->
Disentangling the complexity of drivers and responses in community structure is especially pressing for marine systems. Marine systems have experienced greater changes than other habitats [@antao2020]. Yet, we currently lack linkages between ecological theory, primarily derived from terrestrial systems, and data in marine systems leaving unanswered questions about biodiversity changes in these habitats [@blowes2019]. We propose a framework to explore the nature of biodiversity change in an estuary bay community and connect this change to intrinsic and extrinsic drivers. This research will build towards a more coherent theory of biodiversity change through time [@shoemaker2020].

<!--- --->
Here, we quantified biodiversity changes within a shallow estuarine community over a decade to understand shifts in total species richness, the relative distribution of abundant and rare species, and potential change points in community composition through time...

# Methods

## Sampling area

This study was focused on Terrebonne Bay, located in the northern Gulf of Mexico of Louisiana, USA (29^$\circ$^11.200'N, 90^$\circ$^36.560'W; Figure S1). Terrebonne Bay is a broad (1,761 km2), shallow estuary (average depth ~2m), separated from the Gulf of Mexico by two barrier island chains and connected by four tidal passes. The bay-estuary complex is part of an abandoned distributary of the Mississippi River and therefore no longer receives significant riverine inputs. Thus, the majority of its freshwater inputs are from the local drainage basin [@bianchi2009]. While water column salinity and density are homogeneous, strong salinity gradients exist seaward from the marsh to the barrier islands [@bianchi2009]. Tides in Terrebonne Bay are diurnal with a semidiurnal component, ranging 20-80 cm [@inoue2000].

## Trawl sampling

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's (LUMCON) educational outreach program (https://www.lumcon.edu/education-and-outreach) from 2009 to 2019. Individual trawls were conducted using *trawl apparatus description...* for 10 - 20 min. To dampen potentially volatile trawl-to-trawl variability, we combined all trawls monthly and therefore report changes based on monthly aggregated trawl sampling.

## Community characterization

Species names were checked for typos and confirmed using the *gni_parse* function from the *taxize* package [@chamberlain2020] and further standardized to consistent taxonomic resolution for the entire data series (Supplemental materials). To track trends in species richness through time, we calculated  effort-controlled  species richness, *S~n~*, which provides the expected number of species given a defined number of *n* randomly sampled individuals [@gotelli2001]. For annual and monthly *S~n~* calculations, *n* was set as the minimum number of individuals sampled. We also calculated the uncorrected accumulation of species over time by filtering species from the data set after first occurrence. We then calculated asymptotic species richness, *S~asymp~*, using the Chao1 estimator [@chao1984; @chao1987]. This metric is highly correlated with measured richness, *S* [@mcgill2011], but attempts to extrapolate the total species richness based on incomplete sampling of the community. We calculated *S~asymp~* and *S~n~* using the mobr package [@mcglinn2019; @mcglinn2021a].

Beyond actual and expected numbers of species, we calculated additional measures of biodiversity to capture community change through time. We estimated community turnover, the combination of gains and losses relative to total species richness, and the contributions of species gains and losses using the `codyn` package [@hallett2020]. To quantify changes in the distribution of abundant and rare species, we calculated specific orders of diversity, the percentage of rare species, and evenness in species abundances. An order of diversity, *^q^D*, where *q* represents a non-negative integer, captures the "effective numbers of species" based on varying degrees of rarity of the species in a community [@hill1973]. We used the the *hill_taxa* function in the hillR package [@li2018] to estimate the effective species for diversity order 2 (i.e., *^2^D*). *^2^D* is equivalent to the reciprocal of Simpson's diversity [@simpson1949] (i.e., 1/Simpson's index) which emphasizes the importance of dominant (highly abundant) species. At the other end of the abundance spectrum, we calculated the percentage of rare species in the community as the number of species with abundances below a threshold of $0.05*n$, divided by the total number of species x 100. Here, *n* is the number of individuals in the sample and the value of 0.05 was chosen because it has been shown to capture rarity well and is generally uncorrelated with other metrics of biodiversity. Lastly, to integrate changes in abundance of dominant and rare species within the community, we calculated the evenness of species abundances with the Gini coefficient [@gini1921], normalized for differences in species richness among years, $G^*$ [@solomon1975; @chao2019], :

$$ G^* = (2 \sum_{i = 1}^S ip_i -2)/(S-1)$$

where, in a community of *S* species, species' relative abundances, $p_{i}$, are ordered such that $p_{1} \geq p_{2} \geq ...p_{S}$ for all species of the community. For all metrics, variability was calculated with bootstrapping methods by resampling with replacement `r nperm` times. From these permutations the 2.5% and 97.5% percentiles of yearly statistics were taken to estimate 95% percentile intervals.

We calculated the contribution of each species to community dis/similarity across orders of diversity

## Statistical Analyses

To detect trends in species richness and other dimensions of biodiversity (e.g., turnover, evenness, etc.) we used generalized additive models (GAMs) with a bayesian implementations using the `mgcv` [@wood2011] and  `brms` packages in R [@burkner2017; @burkner2018]. Full models included alinear term for trawl date and smoothing spline terms for month:
$$y = \beta_0 + \beta_1 * time + f_{season}(x_1) + \varepsilon,\;\;\; \varepsilon ~ N(0, \sigma^{2} \Lambda) $$
where, $y$ is the response variable (e.g., richness, turnover, evenness, etc.), $\beta_0$ is an intercept, $\beta_1$ a coefficient for trawl date, and $f_{season}$ represents seasonal smooth functions for month of year. Seasonal terms were fit with cyclic cubic regression splines with smoothing dimension *k* = 12 to allow for continuity between December and January among model years.

# Results

```{r mobr_stats, echo = FALSE, warning=FALSE, message=FALSE, results='hide'}

## Goals estimate the rarified richness through time
## Show changes in evenness and Hill #s
## Look at compositon 
## Show who is contributing
## Show patterns of most important contributors
# estimate the contribution of each species to diversity changes

#### ----

## Annual statistics for first paragraph outline
TB_trawl_sampling_yr = TB_trawl_data %>%
  dplyr::filter(year(Date) %ni% year_removals) %>%
  dplyr::select(Date, year, Common_name, Abundance) %>%
  group_by(year) %>%
  dplyr::summarise(trawl_days = length(unique(Date)),
                   S = length(unique(Common_name)),
                   individuals = sum(Abundance, na.rm = TRUE))

total_richness = length(unique(TB_trawl_taxasite_long$Common_name))

## Annual summaries of richness, etc.
TB_trawl_taxasite_yr <- TB_trawl_data %>%
  # exclude 2007, 2020 & 2021
  dplyr::filter(year %ni% year_removals) %>%
  dplyr::select(year, Common_name, Abundance) %>%
  group_by(year, Common_name) %>%
  summarise(Abundance = sum(Abundance, na.rm = TRUE)) %>%
  na.omit %>% #gropu_by(year) %>%
  # remove trawls with less than 20 individuals 
  # dplyr::filter(sum(Abundance) >=20) %>%
  group_by(year, Common_name) %>%
  pivot_wider(names_from = Common_name, values_from = Abundance, values_fill = list(Abundance = 0)) %>%
  ungroup() %>%
  dplyr::select( year, everything())

## yearly stats
yr_chao1 = TB_trawl_taxasite_yr[,-1] %>%
  apply(., 1, mobr::calc_chao1) %>%
  bind_cols(TB_trawl_taxasite_yr$year,.) %>%
  setNames(., c("year","S_asymp")) %>%
  dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d"))

S_n_yr = min(TB_trawl_sampling_yr$individuals, na.rm = TRUE)

yr_Sn = TB_trawl_taxasite_yr[,-1] %>%
  apply(., 1, function(x) mobr::rarefaction(x, effort = S_n_yr, method = "IBR", extrapolate = TRUE)) %>%
  bind_cols(TB_trawl_taxasite_yr$year,.) %>%
  setNames(., c("year","S_n")) %>%
  dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d"))

 yr_pct_rare =TB_trawl_taxasite_yr[,-1] %>%
  apply(., 1, function(x) pct_rare(x, 0.05)) %>%
  bind_cols(TB_trawl_taxasite_yr$year,.) %>%
  setNames(., c("year","pct_rare")) %>%
  dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d"))

yr_group_stats = left_join(yr_chao1, yr_Sn) %>%
  left_join(yr_pct_rare) %>%
  dplyr::select(Date, year, everything()) %>%
  group_by(Date, year) %>%
  pivot_longer(c(S_n,S_asymp,pct_rare),names_to = "index", values_to = "median" ) %>%
  ungroup
#
TB_trawl_sampling_mth = TB_trawl_taxasite_long %>%
  dplyr::select(Date, date_id, Common_name, Abundance) %>%
  dplyr::filter(Abundance > 0) %>%
  group_by(Date, date_id) %>%
  dplyr::summarise(S = length(unique(Common_name)),
                   individuals = sum(Abundance, na.rm = TRUE))

S_n_mth = min(TB_trawl_sampling_mth$individuals, na.rm = TRUE)
## monthly stats
mth_chao1 = TB_trawl_taxasite[,-c(1,2)] %>%
  apply(., 1, mobr::calc_chao1) %>%
  bind_cols(TB_trawl_taxasite$date_id,.) %>%
  setNames(., c("date_id","S_asymp")) %>%
  left_join(TB_trawl_taxasite[,c(1,2)],.,by = 'date_id')

mth_Sn = TB_trawl_taxasite[,-c(1,2)] %>%
  apply(., 1, function(x) mobr::rarefaction(x, effort = S_n_mth, method = "IBR", extrapolate = TRUE)) %>%
  bind_cols(TB_trawl_taxasite$date_id,.) %>%
  setNames(., c("date_id","S_n")) %>%
  left_join(TB_trawl_taxasite[,c(1,2)],.,by = 'date_id')

mth_pct_rare =TB_trawl_taxasite[,-c(1,2)] %>%
  apply(., 1, function(x) pct_rare(x, 0.05)) %>%
  bind_cols(TB_trawl_taxasite$date_id,.) %>%
  setNames(., c("date_id","pct_rare")) %>%
  left_join(TB_trawl_taxasite[,c(1,2)],.,by = 'date_id')

mth_group_stats = left_join(mth_chao1, mth_Sn) %>%
  left_join(mth_pct_rare) %>%
  dplyr::select(Date, date_id, everything()) %>%
  group_by(Date, date_id) %>%
  pivot_longer(c(S_n,S_asymp, pct_rare),names_to = "index", values_to = "median" ) %>%
  ungroup

mth_even = TB_trawl_taxasite %>%
  dplyr::select(-Date, -date_id) %>%
  apply(., 1, gini_even) %>%
  t() %>%
  data.frame() %>%
  bind_cols(TB_trawl_taxasite[,c('Date','date_id')], .) %>%
  setNames(., c("Date","date_id", "gini_raw", "gini_norm")) %>%
  dplyr::mutate(month = month(Date)) %>%
  # pivot_longer(-Date:-date_id, names_to = "measure", values_to = "median") %>%
  ungroup

# mth_even %>%
#   ggplot(aes(x = Date, y = gini_norm))+
#   geom_point()+
#   geom_line()+
#   scale_y_continuous(name = 'Normalized Gini coefficient', limits = c(0,0.5))+
#   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
#                date_labels = "%Y")

mth_group_stats = mth_group_stats %>%
  bind_rows(mth_even %>%
              pivot_longer(-Date:-date_id, names_to = 'index', values_to = 'median'))

## quantify loss and appearance of species in the timeseries
trawl_mth_turnover = purrr::map(c('appearance','disappearance','total'),
                                ~codyn::turnover(df = TB_trawl_taxasite_long %>% dplyr::select(-Date),
                                                 time.var = 'date_id',
                                                 species.var = 'Common_name',
                                                 abundance.var = 'Abundance',
                                                 metric = .x) %>%
                                  data.frame %>%
                                  dplyr::mutate(date_id = date_id,
                                                variable = .x) %>%
                                  dplyr::rename(value = !!names(.[1])) %>%
                                  bind_rows(data.frame(value = NA_real_, variable =.x))) %>%
  bind_rows %>%
  left_join(TB_trawl_taxasite[,c(1,2)],.,by = 'date_id')

gains = trawl_mth_turnover %>%
  dplyr::filter(variable == "appearance")

losses = trawl_mth_turnover %>%
  dplyr::filter(variable == "disappearance")

total = trawl_mth_turnover %>%
  dplyr::filter(variable == "total")

## add trendline in for A and B

```

```{r evenness, echo = FALSE, warning=FALSE, message=FALSE }
if(rerunEven){
source(here::here("sub-projects/trawl/R/run_evenness_models.R"))
} else if(!all(file.exists(c(here::here("sub-projects/trawl/derived-data/even_null.rds"),
                             here::here("sub-projects/trawl/derived-data/even_s.rds"),
                             here::here("sub-projects/trawl/derived-data/even_full.rds"))))){
  source(here::here("sub-projects/trawl/R/run_evenness_models.R"))
} else{
  even_bayes_full_mod = readRDS(here::here("sub-projects/trawl/derived-data/even_full.rds"))
  even_bayes_s_mod = readRDS(here::here("sub-projects/trawl/derived-data/even_s.rds"))
  even_bayes_null_mod = readRDS(here::here("sub-projects/trawl/derived-data/even_null.rds"))
}

```

<!--- Outline the general patterns annually and  --->
Over the sampling period from `r min(TB_trawl_sampling_yr$year, na.rm = TRUE)` to `r max(TB_trawl_sampling_yr$year, na.rm = TRUE)`, `r length(unique(TB_trawl_data$Date))` trawl-days were performed yielding `r format(signif(sum(TB_trawl_data$Abundance, na.rm = TRUE),3), scientific = TRUE)` individuals. Among years, total effort generally increased, both in number of trawl-days and individuals counted, from a minimum effort of `r min(TB_trawl_sampling_yr$trawl_days)` (`r TB_trawl_sampling_yr %>% slice_min(trawl_days) %>% dplyr::select(year) %>% unlist`) to `r max(TB_trawl_sampling_yr$trawl_days)` (`r TB_trawl_sampling_yr %>% slice_max(trawl_days) %>% dplyr::select(year) %>% unlist`) days and from `r min(TB_trawl_sampling_yr$individuals, na.rm = TRUE)` (`r TB_trawl_sampling_yr %>% slice_min(individuals) %>% dplyr::select(year) %>% unlist`) to `r format(signif(max(TB_trawl_sampling_yr$individuals, na.rm = TRUE),3), scientific = TRUE)` (`r TB_trawl_sampling_yr %>% slice_max(individuals) %>% dplyr::select(year) %>% unlist`) individuals. In total, `r length(unique(TB_trawl_data$Common_name))` unique species were identified and new species were consistently observed throughout the sampling period (Figure S1). Within a calendar year, asymptotic species richness (*S~asymp~*) was relatively stable from 2009 to 2012, reaching a minimum value of `r yr_group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` in `r yr_group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(year)`. Estimated *S~asymp~* increased in 2013 to a series maximum of `r yr_group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` in `r yr_group_stats %>% data.frame %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(year)`. However, given the large differences in effort among years, variability in *S~asymp~* may be partly attributable to sampling effects among years. Therefore, we also calculated rarefied richness at `r S_n_yr` individuals, *S~`r S_n_yr`~*,  for each year. Here, effort-controlled species richness, *S~`r S_n_yr`~*, diverged qualitatively from *S~asymp~* and was generally constant with an apparent slight increase in more recent sampling years. Overall, *S~`r S_n_yr`~* ranged from `r yr_group_stats %>% dplyr::filter(index == "S_n") %>% slice_min(median) %>% dplyr::select(median) %>% round(1)` to `r yr_group_stats %>% dplyr::filter(index == "S_n") %>% slice_max(median) %>% dplyr::select(median) %>% round(1)` among all years. Monthly, of rarefied richness averaged `r round(mean(mth_Sn$S_n, na.rm = TRUE),1)` [range: `r round(min(mth_Sn$S_n, na.rm = TRUE),1)`--`r round(max(mth_Sn$S_n, na.rm = TRUE),1)`] and exhibited seasonal oscillations, but no detectable trend over time (Figure 1A, Table S1).

```{r figure 1, echo = FALSE, warning = FALSE, message=FALSE, fig.height=8.5, fig.width=8.5, fig.cap="Figure 1. A) Rarefied species richness within the nekton community between the years 2009 to 2019 show variable, but consistent, numbers of species over time. B) Despite no trend in the total number of species, the trawl community showed continual turnover throughout."}

## Goals estimate the rarified richness through time
## Show changes in evenness and Hill #s

### create timeseries figure
mth_group_stats %>%
  dplyr::filter(index == 'S_n') %>%
  ggplot(aes( x= Date, y = median))+
  scale_y_continuous(name = expression("Rarefied richness ("~italic(~S[n])~")"), limits = c(1,12),
                     breaks = c(1,5,10),expand = c(0.01,0.01))+
  scale_x_date(date_breaks = "year", limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y",expand = c(0.01,0.01))+
  geom_point(size = 2)+
  geom_line(size = 1.2)+
  theme_tufte()+
  geom_rangeframe(sides = "l")+
  annotate('text', label = "A", x = as.Date("2009-01-30"), y = 11, family ='serif', size = 6, hjust = 0)+
  theme(axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        # panel.grid.minor.x = element_line(color = 'lightgray'),
        plot.margin = unit(c(0.0,0.1,-0.1,0.0), 'cm'), 
        plot.background = element_rect(fill = 'transparent', color = NA),
        panel.background = element_rect(fill = 'transparent', color = NA),
        axis.title.y = element_text(family = 'serif', size = 16),
        axis.text = element_text(family = 'serif', size = 14)) -> rich_time;

trawl_mth_turnover %>%
  dplyr::filter(variable == 'total') %>%
  na.omit %>%
  ggplot(aes(x = Date, y = value, group = variable, color = variable))+
  geom_point(size = 2)+
  geom_line(size = 1.2)+
  scale_color_manual(values = viridis(n = 1, option = 'viridis', begin = .75))+
  scale_y_continuous(name = 'Species turnover',limits = c(0,1), expand = c(0.01,0.01),
                     breaks = c(0,0.2 ,0.4, 0.6, 0.8), labels = c(0,0.2,0.4,0.6,0.8))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y",expand = c(0.01,0.01))+
  theme_tufte()+
  geom_rangeframe(sides = "lb", color = 'black')+
  annotate('text', label = "B", x = as.Date("2009-01-30"), y = 1, family = 'serif', vjust = 1, size = 6, hjust = 0)+
  theme(legend.position = "none",
        legend.justification = c(0,1),
        axis.title.x = element_blank(),
        plot.margin = unit(c(-0.18,0.15,0.0,0.0), 'cm'),
        # panel.grid.minor.x = element_line(color = 'lightgray'),
        plot.background = element_rect(fill = 'transparent', color = NA), 
        axis.title.y = element_text(family = 'serif', size = 16),
        axis.text = element_text(family = 'serif', size = 14)) -> turnover_time

g2 <- ggplotGrob(rich_time)
g3 <- ggplotGrob(turnover_time)
g2$widths <- unit.pmax(g2$widths, g3$widths)
g3$widths <- unit.pmax(g2$widths, g3$widths)
grid.arrange(g2,g3, layout_matrix = rbind(1,1,1,1,2,2))

```

```{r diversity2 plot, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= "Figure 2. A) Temporal trends in the effective number of species of select orders of diversity (q = 0.1, 1,2). Effective species are presented as percentages of ^0^D, or species richness. ^0.1^D highlights the trends in the percentage of species considered very rare, while ^1^D treats rare and dominant species equally, and ^2^D highlights the number of dominant species in the community through time. ^1^D is equivalent to exp(Shannon-Weiner H) and ^2^D is equivalent to 1/Simpson's diversity index. B) The estimated temporal trend for ^q^D orders (q = 0.1--3) from conditional effects of time in generalized additive models. The black line represents the mean estimate surrouned by 25^th^ and 75^th^ (dark grey) and 5^th^ and 95^th^ percentiles (light grey) of the posterior distribution."}

#### patterns of Hill diversity through time}
## hill

hill_time = function(x,...){
  hillR::hill_taxa(TB_trawl_taxasite %>% dplyr::select(-Date,-date_id), q = x) %>%
    data.frame %>% dplyr::mutate(q = paste0("q",as.character(x))) %>% setNames(.,c('value','q')) %>%
    bind_cols(TB_trawl_taxasite %>% dplyr::select(Date, date_id),.)
}
# debugonce(hill_time)
hill_time_df = furrr::future_map(seq(0,3, by = 0.1), ~hill_time(.x)) %>%
  bind_rows %>%
  group_by(Date, date_id) %>% 
  pivot_wider(names_from = 'q', values_from = 'value') %>%
  dplyr::mutate(across(contains('q'), list(~(.x/q0)*100), .names = "{.col}_eff")) %>%
  dplyr::mutate(month = month(Date)) %>%
  dplyr::select(date_id, month, contains('eff'))

eff_cols = names(hill_time_df)[grepl("*eff", names(hill_time_df))]

model_qD = function(x,...){
  form_string = paste(x," ~ date_id + s(month, bs = 'cc', k = 12)")
  bayes_form = bf(formula = formula(form_string))
  mod = brm(bayes_form,
            data = hill_time_df,
            family = gaussian(),
            cores = 4, seed = 42,
            iter = 30000, warmup = 25000, thin = 10,
            control = list(adapt_delta = 0.99),
            save_pars = save_pars(all = TRUE),
            file = paste0(here::here("sub-projects/trawl/derived-data"),"/",as.character(x),"_full.rds"),
            file_refit = 'on_change')

  return(posterior_summary(mod, pars = 'date_id', probs = c(0.025,0.25,0.5,0.75,0.975)))
}

if(rerunD){
  # debugonce(model_qD)
eff_cols %>%
    furrr::future_map(~model_qD(.x)) %>%
    setNames(., nm = eff_cols) %>%
    saveRDS(file = here::here("sub-projects/trawl/derived-data/qD_coefs.rds"))
}

qD_coefs = readRDS( file = here::here("sub-projects/trawl/derived-data/qD_coefs.rds"))
qD_coefs_df = qD_coefs %>% lapply(., data.frame) %>% bind_rows(.id = "model") %>%
  dplyr::mutate(q = as.numeric(sub("^q(\\d{1}|\\d{1}.\\d{1}).*","\\1", model)))

q0.1_bayes_full_mod = readRDS(file = here::here("sub-projects/trawl/derived-data/q0.1_eff_full.rds"))

ceff0.1 <- conditional_effects(q0.1_bayes_full_mod, 'date_id')

ceff0.1_df = data.frame(
  date_id = ceff0.1$date_id$date_id,
  qD = "q0.1_eff",
  estimate = ceff0.1$date_id$estimate__,
  upper = ceff0.1$date_id$upper__,
  lower = ceff0.1$date_id$lower__
)

q0.5_bayes_full_mod = readRDS(file = here::here("sub-projects/trawl/derived-data/q0.5_eff_full.rds"))

ceff0.5 <- conditional_effects(q0.5_bayes_full_mod, 'date_id')

ceff0.5_df = data.frame(
  date_id = ceff0.5$date_id$date_id,
  qD = "q0.5_eff",
  estimate = ceff0.5$date_id$estimate__,
  upper = ceff0.5$date_id$upper__,
  lower = ceff0.5$date_id$lower__
)

q2_bayes_full_mod = readRDS(file = here::here("sub-projects/trawl/derived-data/q2_eff_full.rds"))

ceff2 <- conditional_effects(q2_bayes_full_mod, 'date_id')

ceff2_df = data.frame(
  date_id = ceff2$date_id$date_id,
  qD = "q2_eff",
  estimate = ceff2$date_id$estimate__,
  upper = ceff2$date_id$upper__,
  lower = ceff2$date_id$lower__
)

ceff_full = vctrs::vec_c(ceff0.1_df, ceff0.5_df, ceff2_df) %>%
  dplyr::mutate(Date = as.Date(date_id, origin = "2007-01-01"))

hill_time_df %>%
  dplyr::select(Date, date_id, q0.1_eff, q0.5_eff, q2_eff) %>%
  pivot_longer(-Date:-date_id, names_to = "qD", values_to = "value") %>%
  bind_rows(ceff_full) %>%
  ggplot(aes(x = Date, y = value, group = qD, color = qD))+
  geom_ribbon(data = na.omit(ceff_full), aes(x = Date, ymin = lower, ymax = upper, group = qD, fill = qD),
              color = NA, alpha =0.3, inherit.aes = FALSE)+
  geom_line(data= na.omit(ceff_full), aes(x = Date, y = estimate, group = qD, color = qD), inherit.aes = FALSE)+
  geom_point(size = 1.5,show.legend=FALSE)+
  geom_path(size = 1.2,show.legend=FALSE)+
  guides(fill = 'none')+
  theme_tufte()+
  geom_rangeframe(data = hill_time_df %>%
  dplyr::select(Date, date_id, q0.1_eff, q0.5_eff, q2_eff) %>%
  pivot_longer(-Date:-date_id, names_to = "qD", values_to = "value"), aes(x = Date, y =value), sides = "bl", color = 'black', inherit.aes = FALSE)+
  scale_color_manual(values = viridis(n = 3),labels = c(expression(italic(''^0.1*D)),
                                                        expression(italic(''^1*D)),
                                                        expression(italic(''^2*D)))) +
  scale_fill_manual(values = viridis(n = 3),labels = c(expression(italic(''^0.1*D)),
                                                       expression(italic(''^1*D)),
                                                       expression(italic(''^2*D))))+
  scale_y_continuous(name = expression("Effective species (%)"), limits = c(0,100),
                     expand = c(0.01,0.01)) +
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y", expand = c(0.01,0.01))+
  guides(color = guide_legend(nrow = 1))+
  annotate('text', label = 'A', x = as.Date("2009-01-01"), y = Inf, size = 6, hjust = 0, vjust = 1, family ='serif')+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.background = element_rect(fill = 'transparent', color = NA),
        legend.position = c(0,0),
        legend.justification = c(0,0)) -> hill_time_plot

qD_coefs_df = qD_coefs_df %>%
  dplyr::mutate(across(c("Estimate",matches("Q\\d{1,2}")), ~.x *365))

qD_coefs_df %>% 
  ungroup %>%
  dplyr::filter(q != 0) %>%
  ggplot()+
  geom_ribbon(aes(y = q, xmin = Q2.5, xmax = Q97.5), color = 'lightgrey', alpha = 0.3)+
  geom_ribbon(aes(y = q, xmin = Q25, xmax = Q75), color = 'lightgrey', alpha = 0.5)+
  geom_path(aes(y = q, x = Estimate), size = 1.2)+
  annotate('text', label = 'B', x = -2.5, y = 0, hjust = 0, vjust = 1, size = 6, family = 'serif')+
  theme_tufte()+
  geom_rangeframe(data = qD_coefs_df %>% 
  ungroup %>%
  dplyr::filter(q != 0),
  sides = "br", color = "black")+
  scale_y_continuous(name = expression(italic(""^q*"D")), 
                         trans = scales::reverse_trans(),
                     limits = c(3,0), expand = c(0.01,0.01), position = 'right')+
  scale_x_continuous(name = expression('Temporal trend ('*Delta*'%/year)'),
                     expand = c(0.01,0.01), limits = c(-2.5,0)) -> qD_trend_plot

g4 <- ggplotGrob(hill_time_plot)
g5 <- ggplotGrob(qD_trend_plot)
g4$heights <- unit.pmax(g4$heights, g5$heights)
g5$heights <- unit.pmax(g4$heights, g5$heights)
grid.arrange(g4,g5, layout_matrix = cbind(1,1,1,1,2,2))

even_ceff <- conditional_effects(even_bayes_full_mod, 'date_id')
# plot(ceff)

even_ceff_plot = data.frame(
  date = round(even_ceff$date_id$date_id,0),
  estimate = even_ceff$date_id$estimate__,
  upper = even_ceff$date_id$upper__,
  lower = even_ceff$date_id$lower__
) %>%
  dplyr::mutate(Date = as.Date(date, origin = as.Date("2007-01-01")))

# Add in horizontal lines

```

Consistent with the pattern of continued accumulation of new species throughout the sampling series and relatively stable effort-controlled richness, species gains (`r round(mean(gains$value, na.rm = TRUE),2)` $\pm$ `r round(sd(gains$value, na.rm = TRUE),2)` SD) and losses (`r round(mean(losses$value, na.rm = TRUE),2)` $\pm$ `r round(sd(losses$value, na.rm = TRUE),2)` SD) generally balanced over time. Total turnover was variable but relatively stable (`r round(mean(total$value, na.rm = TRUE),2)` $\pm$ `r round(sd(total$value, na.rm = TRUE),2)` SD) and exhibited no detectable trend or coherent seasonal component (Figure 1B). Species turnover was likely confined to species with moderate relative abundances. While we observed declines in the effective number of species for highly rare (^0.1^D; conditional effects `r round(summary(q0.1_bayes_full_mod)[['fixed']]['date_id','Estimate'],3)`, 95% CI: `r round(summary(q0.1_bayes_full_mod)[['fixed']]['date_id','l-95% CI'],3)`--`r round(summary(q0.1_bayes_full_mod)[['fixed']]['date_id','u-95% CI'],3)`) and highly abundant species (^2^D; `r round(summary(q2_bayes_full_mod)[['fixed']]['date_id','Estimate'],3)`, 95% CI: `r round(summary(q2_bayes_full_mod)[['fixed']]['date_id','l-95% CI'],3)`--`r round(summary(q2_bayes_full_mod)[['fixed']]['date_id','u-95% CI'],3)`), declines were greater for diversity orders describing moderately abundant species (^0.5^D; `r round(summary(q0.5_bayes_full_mod)[['fixed']]['date_id','Estimate'],3)`, 95% CI: `r round(summary(q0.5_bayes_full_mod)[['fixed']]['date_id','l-95% CI'],3)`--`r round(summary(q0.5_bayes_full_mod)[['fixed']]['date_id','u-95% CI'],3)`, Figure 2). These changes in the distribution of relative abundances within the community led to a decrease in the community evenness through time (Figure X). Gini coefficients normalized for species richness varied from `r mth_even %>% dplyr::select(gini_norm) %>% min %>% unlist %>% round(3)` to `r mth_even %>% dplyr::select(gini_norm) %>% max %>% unlist %>% round(3)` and exhibited estimated declines from initial values of `r even_ceff_plot %>% slice_min(date) %>% dplyr::select(estimate) %>% unlist %>% round(3)` (`r even_ceff_plot %>% slice_min(date) %>% dplyr::select(lower) %>% unlist %>% round(3)`--`r even_ceff_plot %>% slice_min(date) %>% dplyr::select(upper) %>% unlist %>% round(3)` 95% CI) to `r even_ceff_plot %>% slice_max(date) %>% dplyr::select(estimate) %>% unlist %>% round(3)` (`r even_ceff_plot %>% slice_max(date) %>% dplyr::select(lower) %>% unlist %>% round(3)`--`r even_ceff_plot %>% slice_max(date) %>% dplyr::select(upper) %>% unlist %>% round(3)` 95% CI) in 2019.

```{r evenness plot, echo=FALSE, warning=FALSE, message=FALSE,fig.cap= "Figure 4. Evenness in species' relative abundances normalized to species richness varied seasonally and exhibited a moderate decrease through time. The blue line represents the mean estimated trend surrounded by the 5^th^ and 95^th^ posterior percentile "}

even_ceff_plot %>%
  ggplot()+
  geom_ribbon(aes(x = Date, ymin = lower, ymax = upper), color = "grey", fill = 'grey', alpha = 0.4)+
  geom_line(aes(x = Date, estimate), color = 'blue', size = 1.1)+
  geom_point(data = mth_even, aes(x = Date, y = gini_norm), color = 'black')+
  geom_line(data = mth_even, aes(x = Date, y = gini_norm), color = 'black', size = 1.2)+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y")+
  scale_y_continuous(name = "Normalized Gini coefficient", limits = c(0,0.55), expand = c(0.01,0.01),
                     breaks = c(0,0.25,0.5))+
  theme_tufte() +
  geom_rangeframe(data = mth_even,aes(x = Date, y = gini_norm),
                  sides = "bl", color = "black", na.rm = TRUE, inherit.aes = FALSE)+
  theme(axis.title.x = element_blank())
```

```{r spp contributions, fig.height=8, fig.width=3.5, dpi=450, fig.cap="Figure 5."}

### create ordering based on relative abundances -------------
TB_trawl_order =  TB_trawl_data %>%
  dplyr::select(Date, Common_name, Abundance) %>%
  dplyr::filter(year(Date) %ni% year_removals) %>%
  group_by(Date) %>%
  dplyr::mutate(rel_abun = Abundance/sum(Abundance, na.rm = TRUE)) %>%
  dplyr::select(-Abundance) %>%
  ungroup %>%
  group_by(Common_name) %>%
  dplyr::summarise(rel_abun = mean(rel_abun, na.rm = TRUE)) %>%
  arrange(desc(rel_abun)) %>%
  na.omit %>% dplyr::filter(!grepl("unspecified", Common_name, ignore.case = TRUE)) %>%
  dplyr::select(Common_name) %>%
  unlist %>% unname

TB_trawl_order_names = fuzzySim::spCodes(TB_trawl_order, nchar.gen = 4, nchar.sp = 6) %>%
  setNames(.,make.names(TB_trawl_order))

TB_trawl_taxasite_dis = TB_trawl_taxasite_long %>%
  dplyr::select(-Date) %>%
  pivot_wider(names_from = 'date_id', values_from = 'Abundance') %>%
  column_to_rownames('Common_name')

TB_trawl_q0.1_dis = dis1(TB_trawl_taxasite_dis, q = 0.1, type = 'tax') %>%
  .[,TB_trawl_order] %>%
  data.frame %>%
  rownames_to_column('dis_type') %>%
  dplyr::filter(dis_type == "UqN") %>%
  pivot_longer(-dis_type, names_to = 'Common_name', values_to = 'value') %>%
  dplyr::mutate(common_label = recode(Common_name, !!!TB_trawl_order_names),
                rel_value = value/sum(value, na.rm = TRUE))

TB_trawl_q0.1_cont = TB_trawl_q0.1_dis %>%
  dplyr::filter(floor(rel_value*100) >=1) %>%
  nrow

TB_trawl_q0.5_dis = dis1(TB_trawl_taxasite_dis, q = 0.5, type = 'tax') %>%
  .[,TB_trawl_order] %>%
  data.frame %>%
  rownames_to_column('dis_type') %>%
  dplyr::filter(dis_type == "UqN") %>%
  pivot_longer(-dis_type, names_to = 'Common_name', values_to = 'value') %>%
  dplyr::mutate(common_label = recode(Common_name, !!!TB_trawl_order_names),
                rel_value = value/sum(value, na.rm = TRUE))

TB_trawl_q0.5_cont = TB_trawl_q0.5_dis %>%
  dplyr::filter(floor(rel_value*100) >=1) %>%
  nrow

TB_trawl_q1_dis = dis1(TB_trawl_taxasite_dis, q = 1, type = 'tax') %>%
  .[,TB_trawl_order] %>%
  data.frame %>%
  rownames_to_column('dis_type') %>%
  dplyr::filter(dis_type == "UqN") %>%
  pivot_longer(-dis_type, names_to = 'Common_name', values_to = 'value') %>%
  dplyr::mutate(common_label = recode(Common_name, !!!TB_trawl_order_names),
                rel_value = value/sum(value, na.rm = TRUE))

TB_trawl_q1_cont = TB_trawl_q1_dis %>%
  dplyr::filter(floor(rel_value*100) >=1) %>%
  nrow

TB_trawl_q2_dis = dis1(TB_trawl_taxasite_dis, q = 2, type = 'tax') %>%
  .[,TB_trawl_order] %>%
  data.frame %>%
  rownames_to_column('dis_type') %>%
  dplyr::filter(dis_type == "UqN") %>%
  pivot_longer(-dis_type, names_to = 'Common_name', values_to = 'value') %>%
  dplyr::mutate(common_label = recode(Common_name, !!!TB_trawl_order_names),
                rel_value = value/sum(value, na.rm = TRUE))

TB_trawl_q2_cont = TB_trawl_q2_dis %>%
  dplyr::filter(floor(rel_value*100) >=1) %>%
  nrow

## Make plots
q0.1_dis_plot = TB_trawl_q0.1_dis %>%
  dplyr::mutate(Common_name = factor(Common_name, levels = make.names(TB_trawl_order)),
                common_label = factor(common_label, levels = TB_trawl_order_names[make.names(TB_trawl_order)])) %>%
  slice_head(n = 20) %>%
  ggplot()+
  geom_col(aes(x= common_label, y = value))+
  scale_y_continuous(name = "Species contribution", limits = c(0,0.06), expand = c(0.0001,0.0001))+
  theme_tufte()+
  geom_rangeframe(sides = "bl", color = "black")+
  annotate("text", label = expression(italic(""^q*"D = 0.1")), x = Inf, y = Inf, vjust = 1, hjust = 1, size = 6, family = "serif")+
  theme(axis.text.x = element_blank(),
        axis.title = element_blank());

q0.5_dis_plot = TB_trawl_q0.5_dis %>%
  dplyr::mutate(Common_name = factor(Common_name, levels = make.names(TB_trawl_order)),
                common_label = factor(common_label, levels = TB_trawl_order_names[make.names(TB_trawl_order)])) %>%
    slice_head(n = 20) %>%
ggplot()+
  geom_col(aes(x= common_label, y = value))+
  scale_y_continuous(name = "Species contribution", limits = c(0,NA), expand = c(0.0001,0.0001))+
  theme_tufte()+
  geom_rangeframe(sides = "bl", color = "black")+
  annotate("text", label = expression(italic(""^q*"D = 0.5")), x = Inf, y = Inf, vjust = 1, hjust = 1, size = 6, family = "serif")+
  theme(axis.text.x = element_blank(),
        axis.title = element_blank());

q1_dis_plot = TB_trawl_q1_dis %>%
  dplyr::mutate(Common_name = factor(Common_name, levels = make.names(TB_trawl_order)),
                common_label = factor(common_label, levels = TB_trawl_order_names[make.names(TB_trawl_order)])) %>%
    slice_head(n = 20) %>%
ggplot()+
  geom_col(aes(x= common_label, y = value))+
  scale_y_continuous(name = "Species contribution", limits = c(0,NA), expand = c(0.0001,0.0001))+
  theme_tufte()+
  geom_rangeframe(sides = "bl", color = "black")+
  annotate("text", label = expression(italic(""^q*"D = 1")), x = Inf, y = Inf, vjust = 1, hjust = 1, size = 6, family = "serif")+
  theme(axis.text.x = element_blank(),
        axis.title = element_blank());

q2_dis_plot = TB_trawl_q2_dis %>%
  dplyr::mutate(Common_name = factor(Common_name, levels = make.names(TB_trawl_order)),
                common_label = factor(common_label, levels = TB_trawl_order_names[make.names(TB_trawl_order)])) %>%
    slice_head(n = 20) %>%
ggplot()+
  geom_col(aes(x= common_label, y = value))+
  scale_y_continuous(name = "Species contribution", limits = c(0,NA), expand = c(0.0001,0.0001))+
  theme_tufte()+
  geom_rangeframe(sides = "bl", color = "black")+
  annotate("text", label = expression(italic(""^q*"D = 2")), x = Inf, y = Inf, vjust = 1, hjust = 1, size = 6, family = "serif")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1),
        axis.title = element_blank())

grid.draw(rbind(ggplotGrob(q0.5_dis_plot),ggplotGrob(q1_dis_plot), ggplotGrob(q2_dis_plot), size = 'last'))

```

```{r spp dynamics, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Figure 5. Scaled abundance of the species with the largest contribution to community dissimilarity within the timeseries at orders of diversity q = 1 (top) and q = 2 (bottom)." }

TB_q2_spp = TB_trawl_q2_dis %>%
  arrange(desc(value)) %>%
 dplyr::select(Common_name) %>%
  unlist %>% .[1:5]

TB_q1_spp = TB_trawl_q1_dis %>%
  arrange(desc(value)) %>%
 dplyr::select(Common_name) %>%
  unlist %>% .[1:5]

TB_q0.5_spp = TB_trawl_q0.5_dis %>%
  arrange(desc(value)) %>%
 dplyr::select(Common_name) %>%
  unlist %>% .[1:5]

TB_q0.1_spp = TB_trawl_q0.1_dis %>%
  arrange(desc(value)) %>%
 dplyr::select(Common_name) %>%
  unlist %>% .[1:5]

TB_q2_spp_df = TB_trawl_taxasite_long %>%
  dplyr::mutate(fixed_common = make.names(Common_name)) %>%
  dplyr::filter(fixed_common %in% TB_q2_spp) %>%
  left_join(TB_trawl_effort_mth, by = 'Date') %>%
  group_by(Common_name) %>%
  dplyr::mutate(corrAbundance = Abundance/trawl_n) %>%
  dplyr::mutate(scale_n = scale(corrAbundance, center = FALSE, scale = TRUE))

TB_q1_spp_df = TB_trawl_taxasite_long %>%
  dplyr::mutate(fixed_common = make.names(Common_name)) %>%
  dplyr::filter(fixed_common %in% TB_q1_spp) %>%
  left_join(TB_trawl_effort_mth, by = 'Date') %>%
  group_by(Common_name) %>%
  dplyr::mutate(corrAbundance = Abundance/trawl_n) %>%
  dplyr::mutate(scale_n = scale(corrAbundance, center = FALSE, scale = TRUE))

TB_q0.5_spp_df = TB_trawl_taxasite_long %>%
  dplyr::mutate(fixed_common = make.names(Common_name)) %>%
  dplyr::filter(fixed_common %in% TB_q0.5_spp) %>%
  left_join(TB_trawl_effort_mth, by = 'Date') %>%
  group_by(Common_name) %>%
  dplyr::mutate(corrAbundance = Abundance/trawl_n) %>%
  dplyr::mutate(scale_n = scale(corrAbundance, center = FALSE, scale = TRUE))

TB_q0.1_spp_df = TB_trawl_taxasite_long %>%
  dplyr::mutate(fixed_common = make.names(Common_name)) %>%
  dplyr::filter(fixed_common %in% TB_q0.1_spp) %>%
  left_join(TB_trawl_effort_mth, by = 'Date') %>%
  group_by(Common_name) %>%
  dplyr::mutate(corrAbundance = Abundance/trawl_n) %>%
  dplyr::mutate(scale_n = scale(corrAbundance, center = FALSE, scale = TRUE))

TB_q2_spp_plot = TB_q2_spp_df %>%
  ggplot()+
  geom_line(aes(x = Date, y = scale_n, group = Common_name, color = Common_name), size = 1.2)+
  scale_color_viridis(discrete = TRUE)+
  scale_y_continuous(name = "Scaled Abundance", expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y", expand = c(0.01,0.01))+
  theme_tufte()+
  geom_rangeframe(data = TB_q2_spp_df, aes(x = Date, y = scale_n),
                  sides = 'bl', color = 'black', inherit.aes = FALSE)+
  annotate('text', label = expression(italic(""^q*"D = 2")), x = as.Date("2019-12-31"), y = Inf, vjust = 1, hjust = 1, size = 6, family = 'serif')+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        legend.position = c(0,1),
        legend.justification = c(0,1),
        legend.background = element_rect(fill = 'transparent', color = NA));#TB_q2_spp_plot

TB_q1_spp_plot = TB_q1_spp_df %>%
  ggplot()+
  geom_line(aes(x = Date, y = scale_n, group = Common_name, color = Common_name), size = 1.2)+
  scale_color_viridis(discrete = TRUE)+
  scale_y_continuous(name = "Scaled Abundance", expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y", expand = c(0.01,0.01))+
  theme_tufte()+
  geom_rangeframe(data = TB_q1_spp_df, aes(x = Date, y = scale_n),
                  sides = 'bl', color = 'black', inherit.aes = FALSE)+
  annotate('text', label = expression(italic(""^q*"D = 1")), x = as.Date("2019-12-31"), y = Inf, vjust = 1, hjust = 1, size = 6, family = 'serif')+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        legend.position = c(0,1),
        legend.justification = c(0,1),
        legend.background = element_rect(fill = 'transparent', color = NA));#TB_q1_spp_plot

TB_q0.5_spp_plot = TB_q0.5_spp_df %>%
  ggplot()+
  geom_line(aes(x = Date, y = scale_n, group = Common_name, color = Common_name), size = 1.2)+
  scale_color_viridis(discrete = TRUE)+
  scale_y_continuous(name = "Scaled Abundance", expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y", expand = c(0.01,0.01))+
  theme_tufte()+
  geom_rangeframe(data = TB_q0.5_spp_df, aes(x = Date, y = scale_n),
                  sides = 'bl', color = 'black', inherit.aes = FALSE)+
  annotate('text', label = expression(italic(""^q*"D = 0.5")), x = as.Date("2019-12-31"), y = Inf, vjust = 1, hjust = 1, size = 6, family = 'serif')+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        legend.position = c(0,1),
        legend.justification = c(0,1),
        legend.background = element_rect(fill = 'transparent', color = NA));#TB_q0.5_spp_plot

TB_q0.1_spp_plot = TB_q0.1_spp_df %>%
  ggplot()+
  geom_line(aes(x = Date, y = scale_n, group = Common_name, color = Common_name))+
  scale_color_viridis(discrete = TRUE)+
  scale_y_continuous(name = "Scaled Abundance", expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y", expand = c(0.01,0.01))+
  theme_tufte()+
  geom_rangeframe(data = TB_q0.1_spp_df, aes(x = Date, y = scale_n),
                  sides = 'bl', color = 'black', inherit.aes = FALSE)+
  annotate('text', label = expression(italic(""^q*"D = 0.1")), x = as.Date("2019-12-31"), y = Inf, vjust = 1, hjust = 1, size = 6, family = 'serif')+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        legend.position = c(0,1),
        legend.justification = c(0,1),
        legend.background = element_rect(fill = 'transparent', color = NA));#TB_q0.1_spp_plot

grid.draw(rbind(ggplotGrob(TB_q0.5_spp_plot), ggplotGrob(TB_q2_spp_plot), size = 'last'))

```

```{r trawl contributions,echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Figure 6. The contribution of individual trawl-dates to dissimilarity of the entire timeseries for diversity order q = 0.5 (top), q = 1 (middle), and q = 2 (bottom)." }

TB_trawl_q0_dis2 = dis1(TB_trawl_taxasite_dis, q = 0, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q0.5_dis2 = dis1(TB_trawl_taxasite_dis, q = 0.5, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q1_dis2 = dis1(TB_trawl_taxasite_dis, q = 1, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q2_dis2 = dis1(TB_trawl_taxasite_dis, q = 2, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q0_dis2 = dis1(TB_trawl_taxasite_dis, q = 0, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q0.1_dis2 = dis1(TB_trawl_taxasite_dis, q = 0.1, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q1_dis2 = dis1(TB_trawl_taxasite_dis, q = 1, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

TB_trawl_q2_dis2 = dis1(TB_trawl_taxasite_dis, q = 2, type = 'tax', type2 = 'k') %>%
  data.frame(check.names = FALSE) %>% 
  rownames_to_column('dis_type') %>%
  pivot_longer(-dis_type,names_to ='date_id', values_to = 'value') %>%
  dplyr::mutate(date_id = as.numeric(date_id)) %>%
  left_join(TB_trawl_taxasite %>% dplyr::select(Date, date_id))

# plot contributions of each trawl to differences
TB_trawl_q0_dis2 %>%
  dplyr::filter(dis_type == "UqN")%>%
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  geom_path()+
  scale_y_continuous(expand = c(0.0001,0.0001))+
   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y")+
  theme_tufte()+
  geom_rangeframe(sides = "lb", color = "black")+
  annotate('text', label = expression(italic(''^q*"D = 0")), x = as.Date("2009-01-01"), y = Inf, size = 6, family = 'serif')+
  theme(axis.title = element_blank(),
        axis.text.x = element_blank())-> TB_q0_dis2_plot

TB_trawl_q0.1_dis2 %>%
  dplyr::filter(dis_type == "UqN")%>%
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  geom_path()+
  scale_y_continuous(expand = c(0.0001,0.0001))+
   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y")+
  theme_tufte()+
  geom_rangeframe(sides = "lb", color = "black")+
  annotate('text', label = expression(italic(''^q*"D = 0.1")), x = as.Date("2009-01-01"), y = Inf, size = 6, family = 'serif')+
  theme(axis.title = element_blank(),
        axis.text.x = element_blank())-> TB_q0.1_dis2_plot

TB_trawl_q0.5_dis2 %>%
  dplyr::filter(dis_type == "UqN")%>%
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  geom_path()+
  scale_y_continuous(expand = c(0.0001,0.0001))+
   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y")+
  theme_tufte()+
  geom_rangeframe(sides = "lb", color = "black")+
  annotate('text', label = expression(italic(''^q*"D = 0.5")), x = as.Date("2019-12-31"), y = Inf, size = 6, family = 'serif', vjust = 1, hjust =1)+
  theme(axis.title = element_blank(),
        axis.text.x = element_blank())-> TB_q0.5_dis2_plot

TB_trawl_q1_dis2 %>%
  dplyr::filter(dis_type == "UqN")%>%
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  geom_path()+
  scale_y_continuous(expand = c(0.0001,0.0001))+
   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y")+
  theme_tufte()+
  geom_rangeframe(sides = "lb", color = "black")+
  annotate('text', label = expression(italic(''^q*"D = 1")), x = as.Date("2019-12-31"), y = Inf, size = 6, family = 'serif', vjust = 1, hjust = 1)+
  theme(axis.title = element_blank(),
        axis.text.x = element_blank())-> TB_q1_dis2_plot

TB_trawl_q2_dis2 %>%
  dplyr::filter(dis_type == "UqN")%>%
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  geom_path()+
  scale_y_continuous(expand = c(0.0001,0.0001))+
   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y")+
  theme_tufte()+
  geom_rangeframe(sides = "lb", color = "black")+
  annotate('text', label = expression(italic(''^q*"D = 2")), x = as.Date("2019-12-31"), y = Inf, size = 6, family = 'serif', vjust = 1, hjust = 1)+
  theme(axis.title= element_blank()) -> TB_q2_dis2_plot

grid.draw(rbind(ggplotGrob(TB_q0.5_dis2_plot),
                ggplotGrob(TB_q1_dis2_plot), 
                ggplotGrob(TB_q2_dis2_plot), size = 'last'))
```

Changes in the estuarine community, in both rare and common species, were not evenly distributed among taxa (Figure 5). For example, dissimilarity in rarer species (i.e., low diversity orders, *^0.1^D*), was contributed by a relatively large number of species (~`r TB_trawl_q0.1_cont` contibuted $\geq$ 1% to total dissimilarity) and individual species effects were relatively diffuse (largest contribution was `r round(max(TB_trawl_q0.1_dis$rel_value*100),1)`). On the contrary, few species contributed to dissimilarity in more abundant species (~`r TB_trawl_q2_cont` contribute $\geq$ 1% to total dissimilarity) and two taxonomic groups contributed `~90% to total dissimilarity (Penaeid spp. and *Micropogonias undulatus*).

Further, community changes were not equally distributed through time (Figure 6). New

# Discussion

# Acknowledgements

We are grateful the LUMCON staff for the logistical support of LUMCON's research, education, and outreach activities allowing the collection of these data. Notably, we extend thanks to the LUMCON vessels staff, crew of the R/V Pelican, the number of LUMCON instructors who supported the safe and successful collection of biological data, and the environmental monitoring department for management of environmental data.. . Trawl sampling was collected under **permit information**...

\newpage

# References

<div id="refs"></div>

# Appendix

```{r map fig S1 }

```

```{r env data fig S2, fig.cap= "Figure S2. Time series of environmental variables a) water temperature and b) salinity measured at the Terrebonne Bay environmental monitoring station (XX N, XX W). Data were provided by the Environmental Monitoring department of LUMCON. Periods of no data in the salinity series represent periods of no data due to storm damage to monitoring stations."}

source(here::here("sub-projects/trawl/R/load_enviroData.R"))

temperature_plot = 
temperature_series %>%
  dplyr::filter(year(Date) %ni% year_removals) %>%
  ggplot()+
  # geom_ribbon(aes(x = Date, ymin = lower, ymax = upper), alpha = 0.5)+
  geom_line(aes(x = Date, y = TB_tempC_series), size = 1)+
  scale_y_continuous(name = expression("Temperature ("~degree*C~")"), limits = c(0,NA),
                     expand = c(0.01,0.01))+
  scale_x_date(date_breaks = "year", limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y",expand = c(0.01,0.01))+
  annotate('text', label = 'A', x = as.Date("2009-01-01"), y = Inf, vjust = 1, hjust = 0, size = 6, family = 'serif')+
  theme_tufte()+
  geom_rangeframe(sides = "bl", color = "black")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())
salinity_plot = 
salinity_series %>%
  ggplot()+
  geom_line(aes(x = Date, y = TB_sal), size = 1, color = 'blue')+
   scale_x_date(date_breaks = "year", limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
               date_labels = "%Y",expand = c(0.01,0.01))+
  scale_y_continuous(name = "Salinity (PSU)", limits = c(0,NA), expand =c(0.01,0.01))+
    annotate('text', label = 'B', x = as.Date("2009-01-01"), y = Inf, vjust = 1, hjust = 0, size = 6, family = 'serif')+
  theme_tufte()+
  geom_rangeframe(sides = "bl", color = "black")+
  theme(axis.title.x = element_blank())

grid.draw(rbind(ggplotGrob(temperature_plot), ggplotGrob(salinity_plot), size = 'last'))
```

```{r supplemental figure s3, fig.cap = "Figure S3."}

# TB_trawl_data %>%
#   # exclude 2020 & 2021
#   dplyr::filter(year %ni% year_removals) %>%
#   dplyr::filter(!duplicated(Common_name )) %>%
#   dplyr::mutate(Date = as.Date(Date)) %>%
#   arrange(Date) %>%
#   dplyr::mutate(richness = 1:n())-> TB_time_accum
# 
# TB_annual_accum <- TB_trawl_data %>%
#   # exclude 2020 & 2021
#   dplyr::filter(year %ni% year_removals) %>%
#   mutate(pres = 1) %>% group_by(year) %>%
#   filter(!duplicated(Common_name)) %>%
#   summarise(richness = sum(pres)) %>%
#   mutate(Date = as.Date(paste(as.character(year),"01","01", sep = "-"), format = "%Y-%m-%d", origin = "2000-01-01"))
# 
# 
# ### create timeseries figure
# TB_time_df <- TB_time_accum %>%
#   dplyr::select(Date, year, richness) %>%
#   mutate(data_type = 'continuous') %>%
#   bind_rows(TB_annual_accum %>% 
#               mutate(data_type = "ann_raw")) %>%
#   bind_rows(group_stats %>% 
#               dplyr::filter(index == "S_n") %>%
#               dplyr::select(Date, year, richness = 'median') %>%
#               dplyr::mutate(data_type = "rarefied")) %>%
#   bind_rows(group_stats %>% 
#               dplyr::filter(index == 'S_asymp') %>%
#               dplyr::select(Date, year, richness = 'median') %>%
#               dplyr::mutate(data_type = "ann_asymp")) %>%
#   dplyr::mutate(data_type = factor(data_type, levels = c("continuous", "rarefied", "ann_raw", "ann_asymp"))) %>%
#   dplyr::filter(data_type == "S_n")
# 
# S_n_lab = paste0("S[",S_n,"]")
# 
# rich_time <-
#   ggplot(TB_time_df, aes(x = Date, y = richness, group = data_type)) +
#   geom_segment(data = TB_time_df %>%
#                  dplyr::filter(data_type %ni% 'continuous') %>%
#                  pivot_wider(Date, names_from = data_type, values_from = richness),
#                aes(x = Date, xend = Date, y = ann_asymp, yend = rarefied ),
#                color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
#                inherit.aes = F) +
#   geom_point(data = TB_time_df, aes(colour = data_type),size = 1.5) +
#   geom_line(data = TB_time_df %>% dplyr::filter(!grepl('ann_asymp', data_type)), aes(colour = data_type),size = 1.2) +
#   scale_x_date(date_breaks = "1 year", date_labels = "%Y",limits = c(as.Date("2009-01-01"),as.Date("2019-12-31"))) +
#   scale_y_continuous(name = "Species Richness")+
#   scale_colour_manual(labels = c("continuous", bquote(italic("S"[.(S_n[1])])),expression(italic("S"["asymp"]))), values = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')) +
#   # scale_alpha_manual(values = c(1,1,0.3))+
#   theme(legend.position = c(0, 1), legend.justification = c(0, 1), legend.background = element_rect(fill = "transparent", colour = NA),
#         legend.title = element_blank(), legend.text.align = 0,
#         # axis.title.x = element_blank(), axis.text.x = element_blank(),
#         legend.spacing.y = unit(-0.1, 'cm'))

```

```{r rarity plot, message=FALSE, warning=FALSE, echo=FALSE}

# group_stats %>%
#   dplyr::filter(index == "pct_rare") %>%
#   dplyr::mutate(year = as.Date(year, format = "%Y")) %>%
#   ggplot()+
#   geom_point(aes(x =Date, y = median), size = 3, shape = 22)+
#   geom_path(aes(x = Date, y = median), size = 1.1)+
#   scale_y_continuous(name = "Percent rare species (%)", limits = c(75 ,100), expand = c(0.01,0.01))+
#   scale_color_manual(values = viridis(n =1, option = 'viridis' ))+
#   scale_fill_manual(values = viridis(n = 1, option = 'viridis'))+
#   scale_x_date(date_breaks = 'year', limits = c(as.Date("2009-01-01"),as.Date("2019-12-31")),
#                date_labels = "%Y")+
#   theme(axis.title.x = element_blank())-> rarity_plot
#   
```

```{r S2 evenness profile, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Figure 5. Species evenness profiles based on Hill numbers relative to annual species richness. ^0^D is equal to the total number of species"}
# TB_trawl_yr_evenness = TB_trawl_taxasite_yr %>%
#   split(., seq(1:nrow(.))) %>%
#   purrr::map(~new_fun(.x, seq(0,4,0.1), "E3")) %>%
#   setNames(., nm = TB_trawl_taxasite_yr$year) %>%
#   purrr::map(function(a) a %>% t %>% data.frame %>%
#                setNames(., nm = seq(0,4,0.1))) %>%
#           bind_rows(.id = 'year') %>%
#   pivot_longer(-year, names_to = 'q', values_to = 'eff_spp') %>%
#   dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d")) %>%
#   dplyr::mutate(across(q, as.numeric))
# 
# TB_trawl_q1_plot =
#   TB_trawl_yr_evenness %>%
#   dplyr::filter(q == 1) %>%
#   ggplot()+
#   geom_line(aes(x = Date, y = eff_spp))+
#   scale_y_continuous(name = 'Eff. Species (%)', limits = c(0,0.2))+
#   scale_x_date(breaks = as.Date(c("2009-01-01","2014-01-01","2019-01-01"),format = "%Y-%m-%d"),
#                date_labels = "%Y")+
#   theme(axis.title.x = element_blank())+
#   annotate('text', x = as.Date(Inf), y = as.Date(Inf), label = expression(italic(""^1*D)), size =5,
#            hjust = 1, vjust = 1)
# 
# TB_trawl_q2_plot =
#   TB_trawl_yr_evenness %>%
#   dplyr::filter(q == 2) %>%
#   ggplot()+
#   geom_line(aes(x = Date, y = eff_spp))+
#   scale_y_continuous(name = 'Eff. Species (%)', limits = c(0,0.15))+
#   scale_x_date(breaks = as.Date(c("2009-01-01","2014-01-01","2019-01-01"),format = "%Y-%m-%d"),
#                date_labels = "%Y")+
#   theme(axis.title.x = element_blank(),
#         axis.title.y = element_blank())+
#   annotate('text', x = as.Date(Inf), y = as.Date(Inf), label = expression(italic(""^2*D)), size =5,
#            hjust = 1, vjust = 1)
# 
# TB_trawl_yr_evenness %>%
#   ggplot() +
#   geom_path(aes(x = q, y = eff_spp, group = year, color = year), size = 1.1)+
#   scale_y_continuous(name = 'Effective species (%)', limits = c(0,1), expand = c(0.01,0.01))+
#   scale_x_continuous(name = expression(italic(""^q*D)), limits = c(0,3), expand = c(0.02,0.02)) +
#   scale_color_manual(values = viridis(n = 12, option = 'viridis')) +
#   annotation_custom(
#     ggplotGrob(TB_trawl_q1_plot),
#     xmin = 0.75, xmax = 1.9, ymin = 0.5, ymax = 1
#   )+
#   annotation_custom(
#     ggplotGrob(TB_trawl_q2_plot),
#     xmin = 2, xmax = 3, ymin = 0.5, ymax = 1
#   )+
#   theme(legend.title = element_blank(),
#         legend.position = 'top')+
#   guides(color=guide_legend(nrow=2,byrow=TRUE))

```